To read file and assign variable
```{r}
library(readr)
titanic <- read_csv(file = "train.csv")
nrow(titanic)
```
Q1) Summarize the population
```{r}
titanic%>%
summarise(mu=mean(PassengerId),pop_med=median(PassengerId), 
sigma = sd(PassengerId), 
pop_iqr = IQR(PassengerId),
pop_min=min(PassengerId),
pop_max=max(PassengerId),
pop_q1=quantile(PassengerId,0.25),
pop_q3=quantile(PassengerId,0.75))
```
Q2) The difference between quantitative and qualitative that: 
Quantitative is expressed in numbers and graph, It used to test,or confirm, theories ad assumptions.
While Qualitative is expressed in words , It used to understand concepts ,or thoughts.
Quantitative: 1.PassengerId 2.Survived 3.Pclass 4.Age 5.SibSp 6.Parch 7.Fare
Qualitative: 1.Name 2.Sex 3.Ticket 4.Cabin 5.Embarked

Q3)i)
```{r}
#To count missing variables of Age
sum(is.na(titanic$Age))
```
ii) 
iii) 
iv)
Q4)Remove missing data
```{r}
titanic=na.omit(titanic)
nrow(titanic)
```

Q5)Draw histogram for dataset's Age
```{r}
hist(titanic$Age , xlab = "Age")
```